---
title: "br_marginal"
author: "Fernando Rawicz"
date: "12/05/2021"
output: html_document
---

```{r message = FALSE}
rm(list = ls())
set.seed(1234)
options(scipen = 9999)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

library(rstan)
library(shinystan)
library(stringr)
library(dplyr)
library(readxl)
library(gtools)
library(rjags)
library(runjags)
library(brms)
library(knitr)
library(DT)
library(tidyverse)
library(bsts)
library(MARSS)
library(reshape2)
library(tidybayes)
library(lubridate)
library(gridExtra)
```


## The data 

The data is the evolution of the polls for all the candidates in Brazilian election of 2018. The polls used are only the ones made by Ibope and Datafolha. Every graph with the evolution is shown below:


```{r message = FALSE}
br_polls <- read_excel("data/pollsbrazil.xltx") %>% 
  arrange(Date) %>% 
  mutate(sumlessnull = ( `PSL`+`PT`+`PDT`+`PSDB`+`REDE`+`NOVO`+
                           `PODE`+`MDB`+`PATRI`+`PSOL`+`PSTU`+`DC`+`PPL` ) ) %>% 
  filter(sumlessnull< 100) %>% 
  filter(Company == "Datafolha" | Company == "Ibope"| Company == "Ipespe")

# DT::datatable(br_polls)
```

```{r message = FALSE}

  
  br_polls2 <- br_polls %>% 
     as.data.frame() %>% 
    mutate(
      PSL = round((PSL *100) / sumlessnull, 2),
      PT =  round((PT *100) / sumlessnull, 2),
      PDT =  round((PDT *100) / sumlessnull, 2),
      PSDB =  round((PSDB *100) / sumlessnull, 2),
      REDE =  round((REDE *100) / sumlessnull, 2),
      Others =  round((( NOVO + PODE + MDB + PATRI + PSOL + PSTU + DC + PPL)*100)/sumlessnull, 2)
    ) %>% 
    dplyr::select(c( "Date", "PSL","PT","PDT","PSDB","REDE", "Others"))


par(mfrow=c(3,2))


plot.xts(xts(x = br_polls2$PSL, order.by = as.Date(br_polls$Date)), main = "PSL" )
 
plot.xts(xts(x = br_polls2$PT, order.by = as.Date(br_polls$Date)), main = "PT" )

plot.xts(xts(x = br_polls2$PDT, order.by = as.Date(br_polls$Date)), main = "PDT" )
  
plot.xts(xts(x = br_polls2$PSDB, order.by = as.Date(br_polls$Date)), main = "PSDB" )
 
plot.xts(xts(x = br_polls2$REDE, order.by = as.Date(br_polls$Date)), main = "REDE" )

plot.xts(xts(x = br_polls2$Others, order.by = as.Date(br_polls$Date)), main = "Others" )




  
```

## Univariate space state model

$$
x_t = x_{t-1} + u +w_t, w_t \sim N(0,q)
$$

$$
y_t = x_{t} + v_t, v_t \sim N(0,r)
$$

```{r message = FALSE}

jagsscriptss <- "
model {  
   # priors on parameters
   u ~ dnorm(0, 1); 
   inv.q ~ dgamma(1, 3); 
   q <- 1/inv.q;
   inv.r ~ dgamma(mu0, mu1);
   r <- 1/inv.r; 
   X0 ~ dnorm(Y1, 0.1);
   
   X[1] ~ dnorm(X0 + u, inv.q);
   EY[1] <- X[1];
   Y[1] ~ dnorm(EY[1], inv.r);
   for(i in 2:N) {
      X[i] ~ dnorm(X[i-1] + u, inv.q);
      EY[i] <- X[i];
      Y[i] ~ dnorm(EY[i], inv.r); 
   }
}  
"
jags.paramsss <- c("q", "r", "EY", "u")

jags.datass.psl <- list(Y = as.vector(br_polls2$PSL), N = length(as.vector(br_polls2$PSL)), Y1 = as.vector(br_polls2$PSL)[1], mu0 = 1.5, mu1 = 6)
jags.datass.pt <- list(Y = as.vector(br_polls2$PT), N = length(as.vector(br_polls2$PT)), Y1 = as.vector(br_polls2$PT)[1], mu0 = 1.5, mu1 = 6)
jags.datass.pdt <- list(Y = as.vector(br_polls2$PDT), N = length(as.vector(br_polls2$PDT)), Y1 = as.vector(br_polls2$PDT)[1], mu0 = 1.5, mu1 = 6)
jags.datass.psdb <- list(Y = as.vector(br_polls2$PSDB), N = length(as.vector(br_polls2$PSDB)), Y1 = as.vector(br_polls2$PSDB)[1], mu0 = 1.5, mu1 = 6)
jags.datass.rede <- list(Y = as.vector(br_polls2$REDE), N = length(as.vector(br_polls2$REDE)), Y1 = as.vector(br_polls2$REDE)[1], mu0 = 1.5, mu1 = 6)
jags.datass.others <- list(Y = as.vector(br_polls2$Others), N = length(as.vector(br_polls2$Others)), Y1 = as.vector(br_polls2$Others)[1], mu0 = 1.5, mu1 = 6)

```

The convergence analysis for this model

```{r message = FALSE}

modss.psl = jags.model(textConnection(jagsscriptss), data=jags.datass.psl, n.chains=3)
update(modss.psl, 1e3)
mod_simss.psl = coda.samples(model=modss.psl, variable.names=jags.paramsss, n.iter=5e3)

modss.pt = jags.model(textConnection(jagsscriptss), data=jags.datass.pt, n.chains=3)
update(modss.pt, 1e3)
mod_simss.pt = coda.samples(model=modss.pt, variable.names=jags.paramsss, n.iter=5e3)

modss.pdt = jags.model(textConnection(jagsscriptss), data=jags.datass.pdt, n.chains=3)
update(modss.pdt, 1e3)
mod_simss.pdt = coda.samples(model=modss.pdt, variable.names=jags.paramsss, n.iter=5e3)

modss.psdb = jags.model(textConnection(jagsscriptss), data=jags.datass.psdb, n.chains=3)
update(modss.psdb, 1e3)
mod_simss.psdb = coda.samples(model=modss.psdb, variable.names=jags.paramsss, n.iter=5e3)

modss.rede = jags.model(textConnection(jagsscriptss), data=jags.datass.rede, n.chains=3)
update(modss.rede, 1e3)
mod_simss.rede = coda.samples(model=modss.rede, variable.names=jags.paramsss, n.iter=5e3)

modss.other = jags.model(textConnection(jagsscriptss), data=jags.datass.others, n.chains=3)
update(modss.other, 1e3)
mod_simss.other = coda.samples(model=modss.other, variable.names=jags.paramsss, n.iter=5e3)


mcmcresultspsl = as.mcmc(do.call(rbind, mod_simss.psl))
mcmcresultspt = as.mcmc(do.call(rbind, mod_simss.pt))
mcmcresultspdt = as.mcmc(do.call(rbind, mod_simss.pdt))
mcmcresultspsdb = as.mcmc(do.call(rbind, mod_simss.psdb))
mcmcresultsrede = as.mcmc(do.call(rbind, mod_simss.rede))
mcmcresultsother = as.mcmc(do.call(rbind, mod_simss.other))

(colMeans(round(mcmcresultspsl,2)))
(colMeans(round(mcmcresultspt,2)))
(colMeans(round(mcmcresultspdt,2)))
(colMeans(round(mcmcresultspsdb,2)))
(colMeans(round(mcmcresultsrede,2)))
(colMeans(round(mcmcresultsother,2)))

```

The summary of the model

```{r message = FALSE}


# summary(mod_simss.psl)
# summary(mod_simss.pt)
# summary(mod_simss.pdt)
# summary(mod_simss.psdb)
# summary(mod_simss.rede)
# summary(mod_simss.other)

```
## Forecasting


Since the last poll is one day before the election, we will treat the election as T + 1. For this we will use a jags model with the last data being NA and the model will forecast for this value.

### Forecast for the Space State model

```{r message = FALSE}
datevec <- c(br_polls2$Date, max(br_polls2$Date) + days(1), max(br_polls2$Date) + days(2))
jags.datass.f.psl <- list(Y = as.vector(c(br_polls2$PSL, NA, NA)), N = length(as.vector(br_polls2$PSL)) + 2, Y1 = as.vector(br_polls2$PSL)[1], mu0 = 1.5, mu1 = 6)
jags.datass.f.pt <- list(Y = as.vector(c(br_polls2$PT, NA, NA)), N = length(as.vector(br_polls2$PT)) + 2, Y1 = as.vector(br_polls2$PT)[1], mu0 = 1.5, mu1 = 6)
jags.datass.f.pdt <- list(Y = as.vector(c(br_polls2$PDT, NA, NA)), N = length(as.vector(br_polls2$PDT)) + 2, Y1 = as.vector(br_polls2$PDT)[1], mu0 = 1.5, mu1 = 6)
jags.datass.f.psdb <- list(Y = as.vector(c(br_polls2$PSDB, NA, NA)), N = length(as.vector(br_polls2$PSDB)) + 2, Y1 = as.vector(br_polls2$PSDB)[1], mu0 = 1.5, mu1 = 6)
jags.datass.f.rede <- list(Y = as.vector(c(br_polls2$REDE, NA, NA)), N = length(as.vector(br_polls2$REDE)) + 2, Y1 = as.vector(br_polls2$REDE)[1], mu0 = 1.5, mu1 = 6)
jags.datass.f.others <- list(Y = as.vector(c(br_polls2$Others, NA, NA)), N = length(as.vector(br_polls2$Others)) + 2, Y1 = as.vector(br_polls2$Others)[1], mu0 = 1.5, mu1 = 6)


jags.paramsss <- c("EY")

modss.f.psl = jags.model(textConnection(jagsscriptss), data=jags.datass.f.psl, n.chains=3)
update(modss.f.psl, 1e3)
mod_simss.f.psl = coda.samples(model=modss.f.psl, variable.names=jags.paramsss, n.iter=5e3)

modss.f.pt = jags.model(textConnection(jagsscriptss), data=jags.datass.f.pt, n.chains=3)
update(modss.f.pt, 1e3)
mod_simss.f.pt = coda.samples(model=modss.f.pt, variable.names=jags.paramsss, n.iter=5e3)

modss.f.pdt = jags.model(textConnection(jagsscriptss), data=jags.datass.f.pdt, n.chains=3)
update(modss.f.pdt, 1e3)
mod_simss.f.pdt = coda.samples(model=modss.f.pdt, variable.names=jags.paramsss, n.iter=5e3)

modss.f.psdb = jags.model(textConnection(jagsscriptss), data=jags.datass.f.psdb, n.chains=3)
update(modss.f.psdb, 1e3)
mod_simss.f.psdb = coda.samples(model=modss.f.psdb, variable.names=jags.paramsss, n.iter=5e3)

modss.f.rede = jags.model(textConnection(jagsscriptss), data=jags.datass.f.rede, n.chains=3)
update(modss.f.rede, 1e3)
mod_simss.f.rede = coda.samples(model=modss.f.rede, variable.names=jags.paramsss, n.iter=5e3)

modss.f.others = jags.model(textConnection(jagsscriptss), data=jags.datass.f.others, n.chains=3)
update(modss.f.others, 1e3)
mod_simss.f.others = coda.samples(model=modss.f.others, variable.names=jags.paramsss, n.iter=5e3)



sumpsl <- summary(mod_simss.f.psl)
sumpt <- summary(mod_simss.f.pt)
sumpdt <- summary(mod_simss.f.pdt)
sumpsdb <- summary(mod_simss.f.psdb)
sumrede <- summary(mod_simss.f.rede)
sumother <- summary(mod_simss.f.others)


```


```{r message = FALSE}

quantilespsl <- sumpsl$quantiles  
ggpsl <- sumpsl$statistics %>% data.frame() %>% dplyr::select("Mean") %>% cbind(datevec) %>% cbind(quantilespsl)

quantilespt <- sumpt$quantiles  
ggpt <- sumpt$statistics %>% data.frame() %>% dplyr::select("Mean") %>% cbind(datevec) %>% cbind(quantilespt)

quantilespdt <- sumpdt$quantiles  
ggpdt <- sumpdt$statistics %>% data.frame() %>% dplyr::select("Mean") %>% cbind(datevec) %>% cbind(quantilespdt)

quantilespsdb <- sumpsdb$quantiles  
ggpsdb <- sumpsdb$statistics %>% data.frame() %>% dplyr::select("Mean") %>% cbind(datevec) %>% cbind(quantilespsdb)

quantilesrede <- sumrede$quantiles  
ggrede <- sumrede$statistics %>% data.frame() %>% dplyr::select("Mean") %>% cbind(datevec) %>% cbind(quantilesrede)

quantilesothers <- sumother$quantiles  
ggothers <- sumother$statistics %>% data.frame() %>% dplyr::select("Mean") %>% cbind(datevec) %>% cbind(quantilesothers)


p1 <- ggplot(ggpsl, aes(datevec)) +
   geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey70") +
   geom_line(aes(y = Mean)) +
   labs(x = "Date",y = "%", title = "PSL Valid Vote Share in %")+
  theme(plot.title = element_text(hjust = 0.5))+ 
   geom_vline(xintercept = as.numeric(ggpsl$datevec[c(26)]) )+
  geom_point(aes(x= ggpsl$datevec[c(27)], y=45.03), colour="red", size =3)



p2 <- ggplot(ggpt, aes(datevec)) +
   geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey70") +
   geom_line(aes(y = Mean)) +
   labs(x = "Date",y = "%", title = "PT Valid Vote Share in %")+
  theme(plot.title = element_text(hjust = 0.5))+ 
   geom_vline(xintercept = as.numeric(ggpt$datevec[c(26)]) )+
  geom_point(aes(x= ggpt$datevec[c(27)], y=29.28), colour="red", size =3)



p3 <- ggplot(ggpdt, aes(datevec)) +
   geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey70") +
   geom_line(aes(y = Mean)) +
   labs(x = "Date",y = "%", title = "PDT Valid Vote Share in %")+
  theme(plot.title = element_text(hjust = 0.5))+ 
   geom_vline(xintercept = as.numeric(ggpdt$datevec[c(26)]) )+
  geom_point(aes(x= ggpdt$datevec[c(27)], y=12.47), colour="red", size =3)



p4 <- ggplot(ggpsdb, aes(datevec)) +
   geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey70") +
   geom_line(aes(y = Mean)) +
   labs(x = "Date",y = "%", title= "PSDB Valid Vote Share in %")+
  theme(plot.title = element_text(hjust = 0.5))+ 
   geom_vline(xintercept = as.numeric(ggpsdb$datevec[c(26)]) )+
  geom_point(aes(x= ggpsdb$datevec[c(27)], y=4.76), colour="red", size =3)



p5 <- ggplot(ggrede, aes(datevec)) +
   geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey70") +
   geom_line(aes(y = Mean)) +
   labs(x = "Date",y = "%", title = "REDE Valid Vote Share in %")+
  theme(plot.title = element_text(hjust = 0.5))+ 
   geom_vline(xintercept = as.numeric(ggrede$datevec[c(26)]) )+
  geom_point(aes(x= ggrede$datevec[c(27)], y=1), colour="red", size =3)



p6 <- ggplot(ggothers, aes(datevec)) +
   geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey70") +
   geom_line(aes(y = Mean)) +
   labs(x = "Date", title = "Other parties Valid Vote Share in %", y = "%")+
  theme(plot.title = element_text(hjust = 0.5))+ 
   geom_vline(xintercept = as.numeric(ggothers$datevec[c(26)]) )+
  geom_point(aes(x= ggothers$datevec[c(27)], y=4.46), colour="red", size =3)




grid.arrange(p1, p2,p3,p4,p5,p6, nrow = 2)

```


## Forecast 10 days




```{r message = FALSE}

br_polls10days <- br_polls2 %>% 
   filter(Date < "2018-09-25")

datevec10 <- c(br_polls10days$Date, max(br_polls10days$Date) + days(1), max(br_polls2$Date) + days(12))


jags.datass.f.psl10 <- list(Y = as.vector(c(br_polls10days$PSL, rep(NA, 12))), N = length(as.vector(br_polls10days$PSL)) + 12, Y1 = as.vector(br_polls10days$PSL)[1], mu0 = 1.5, mu1 = 6)

jags.datass.f.pt10 <- list(Y = as.vector(c(br_polls10days$PT, rep(NA, 12))), N = length(as.vector(br_polls10days$PT)) + 12, Y1 = as.vector(br_polls10days$PT)[1], mu0 = 1.5, mu1 = 6)

jags.datass.f.pdt10 <- list(Y = as.vector(c(br_polls10days$PDT, rep(NA, 12))), N = length(as.vector(br_polls10days$PDT)) + 12, Y1 = as.vector(br_polls10days$PDT)[1], mu0 = 1.5, mu1 = 6)

jags.datass.f.psdb10 <- list(Y = as.vector(c(br_polls10days$PSDB, rep(NA, 12))), N = length(as.vector(br_polls10days$PSDB)) + 12, Y1 = as.vector(br_polls10days$PSDB)[1], mu0 = 1.5, mu1 = 6)

jags.datass.f.rede10 <- list(Y = as.vector(c(br_polls10days$REDE, rep(NA, 12))), N = length(as.vector(br_polls10days$REDE)) + 12, Y1 = as.vector(br_polls10days$REDE)[1], mu0 = 1.5, mu1 = 6)

jags.datass.f.others10 <- list(Y = as.vector(c(br_polls10days$Others, rep(NA, 12))), N = length(as.vector(br_polls10days$Others)) + 12, Y1 = as.vector(br_polls10days$Others)[1], mu0 = 1.5, mu1 = 6)


jags.paramsss <- c("EY")

modss.f.psl10 = jags.model(textConnection(jagsscriptss), data=jags.datass.f.psl10, n.chains=3)
update(modss.f.psl10, 1e3)
mod_simss.f.psl10 = coda.samples(model=modss.f.psl10, variable.names=jags.paramsss, n.iter=5e3)

modss.f.pt10 = jags.model(textConnection(jagsscriptss), data=jags.datass.f.pt10, n.chains=3)
update(modss.f.pt10, 1e3)
mod_simss.f.pt10 = coda.samples(model=modss.f.pt10, variable.names=jags.paramsss, n.iter=5e3)

modss.f.pdt10 = jags.model(textConnection(jagsscriptss), data=jags.datass.f.pdt10, n.chains=3)
update(modss.f.pdt10, 1e3)
mod_simss.f.pdt10 = coda.samples(model=modss.f.pdt10, variable.names=jags.paramsss, n.iter=5e3)

modss.f.psdb10 = jags.model(textConnection(jagsscriptss), data=jags.datass.f.psdb10, n.chains=3)
update(modss.f.psdb10, 1e3)
mod_simss.f.psdb10 = coda.samples(model=modss.f.psdb10, variable.names=jags.paramsss, n.iter=5e3)

modss.f.rede10 = jags.model(textConnection(jagsscriptss), data=jags.datass.f.rede10, n.chains=3)
update(modss.f.rede10, 1e3)
mod_simss.f.rede10 = coda.samples(model=modss.f.rede10, variable.names=jags.paramsss, n.iter=5e3)

modss.f.others10 = jags.model(textConnection(jagsscriptss), data=jags.datass.f.others10, n.chains=3)
update(modss.f.others10, 1e3)
mod_simss.f.others10 = coda.samples(model=modss.f.others10, variable.names=jags.paramsss, n.iter=5e3)



sumpsl10 <- summary(mod_simss.f.psl10)
sumpt10 <- summary(mod_simss.f.pt10)
sumpdt10 <- summary(mod_simss.f.pdt10)
sumpsdb10 <- summary(mod_simss.f.psdb10)
sumrede10 <- summary(mod_simss.f.rede10)
sumothe10r <- summary(mod_simss.f.others10)


```






## Posterior Analysis

```{r}

library(bayesplot)

p11 <- mcmc_areas(
  mcmcresultspsl ,    
  prob = 0.90,
regex_pars = c("EY")
  ) + 
   labs(title="PSL",
        x ="Value in P.P", y = "Parameter")

p12 <- mcmc_areas(
  mcmcresultspt  ,    
  prob = 0.90,
regex_pars = c("EY")
  ) + 
   labs(title="PT",
        x ="Value in P.P", y = "Parameter")

p13 <- mcmc_areas(
  mcmcresultspdt  ,    
  prob = 0.90,
regex_pars = c("EY")
  ) + 
   labs(title="PDT",
        x ="Value in P.P", y = "Parameter")


p15 <-mcmc_areas(
  mcmcresultspsdb,    
  prob = 0.90,
regex_pars = c("EY")
  ) + 
   labs(title="PSDB",
        x ="Value in P.P", y = "Parameter")

p14 <-mcmc_areas(
  mcmcresultsrede,    
  prob = 0.90,
regex_pars = c("EY")
  ) + 
   labs(title="REDE",
        x ="Value in P.P", y = "Parameter")

p16 <-mcmc_areas(
  mcmcresultsother,    
  prob = 0.90,
regex_pars = c("EY")
  ) + 
   labs(title="Others",
        x ="Value in P.P", y = "Parameter")


grid.arrange(p11, p12,p13,p14,p15,p16, nrow = 3)

```





## Sensitivity Analysis




```{r message = FALSE}

jagsscriptss <- "
model {  
   # priors on parameters
   u ~ dnorm(0, 1); 
   inv.q ~ dgamma(1, 3); 
   q <- 1/inv.q;
   inv.r ~ dgamma(1.5, 15);
   r <- 1/inv.r; 
   X0 ~ dnorm(Y1, 0.1);
   
   X[1] ~ dnorm(X0 + u, inv.q);
   EY[1] <- X[1];
   Y[1] ~ dnorm(EY[1], inv.r);
   for(i in 2:N) {
      X[i] ~ dnorm(X[i-1] + u, inv.q);
      EY[i] <- X[i];
      Y[i] ~ dnorm(EY[i], inv.r); 
   }
}  
"
jags.paramsss <- c("q", "r", "EY", "u")

jags.datass.psl <- list(Y = as.vector(br_polls2$PSL), N = length(as.vector(br_polls2$PSL)), Y1 = as.vector(br_polls2$PSL)[1])
jags.datass.pt <- list(Y = as.vector(br_polls2$PT), N = length(as.vector(br_polls2$PT)), Y1 = as.vector(br_polls2$PT)[1])
jags.datass.pdt <- list(Y = as.vector(br_polls2$PDT), N = length(as.vector(br_polls2$PDT)), Y1 = as.vector(br_polls2$PDT)[1])
jags.datass.psdb <- list(Y = as.vector(br_polls2$PSDB), N = length(as.vector(br_polls2$PSDB)), Y1 = as.vector(br_polls2$PSDB)[1])
jags.datass.rede <- list(Y = as.vector(br_polls2$REDE), N = length(as.vector(br_polls2$REDE)), Y1 = as.vector(br_polls2$REDE)[1])
jags.datass.others <- list(Y = as.vector(br_polls2$Others), N = length(as.vector(br_polls2$Others)), Y1 = as.vector(br_polls2$Others)[1])

modss.psl = jags.model(textConnection(jagsscriptss), data=jags.datass.psl, n.chains=3)
update(modss.psl, 1e3)
mod_simss.psl = coda.samples(model=modss.psl, variable.names=jags.paramsss, n.iter=5e3)

modss.pt = jags.model(textConnection(jagsscriptss), data=jags.datass.pt, n.chains=3)
update(modss.pt, 1e3)
mod_simss.pt = coda.samples(model=modss.pt, variable.names=jags.paramsss, n.iter=5e3)

modss.pdt = jags.model(textConnection(jagsscriptss), data=jags.datass.pdt, n.chains=3)
update(modss.pdt, 1e3)
mod_simss.pdt = coda.samples(model=modss.pdt, variable.names=jags.paramsss, n.iter=5e3)

modss.psdb = jags.model(textConnection(jagsscriptss), data=jags.datass.psdb, n.chains=3)
update(modss.psdb, 1e3)
mod_simss.psdb = coda.samples(model=modss.psdb, variable.names=jags.paramsss, n.iter=5e3)

modss.rede = jags.model(textConnection(jagsscriptss), data=jags.datass.rede, n.chains=3)
update(modss.rede, 1e3)
mod_simss.rede = coda.samples(model=modss.rede, variable.names=jags.paramsss, n.iter=5e3)

modss.other = jags.model(textConnection(jagsscriptss), data=jags.datass.others, n.chains=3)
update(modss.other, 1e3)
mod_simss.other = coda.samples(model=modss.other, variable.names=jags.paramsss, n.iter=5e3)


mcmcresultspsl = as.mcmc(do.call(rbind, mod_simss.psl))
mcmcresultspt = as.mcmc(do.call(rbind, mod_simss.pt))
mcmcresultspdt = as.mcmc(do.call(rbind, mod_simss.pdt))
mcmcresultspsdb = as.mcmc(do.call(rbind, mod_simss.psdb))
mcmcresultsrede = as.mcmc(do.call(rbind, mod_simss.rede))
mcmcresultsother = as.mcmc(do.call(rbind, mod_simss.other))

(colMeans(round(mcmcresultspsl,2)))
(colMeans(round(mcmcresultspt,2)))
(colMeans(round(mcmcresultspdt,2)))
(colMeans(round(mcmcresultspsdb,2)))
(colMeans(round(mcmcresultsrede,2)))
(colMeans(round(mcmcresultsother,2)))

```




```{r message = FALSE}

jagsscriptss <- "
model {  
   # priors on parameters
   u ~ dnorm(0, 1); 
   inv.q ~ dgamma(1, 3); 
   q <- 1/inv.q;
   inv.r ~ dgamma(1.5, 30);
   r <- 1/inv.r; 
   X0 ~ dnorm(Y1, 0.1);
   
   X[1] ~ dnorm(X0 + u, inv.q);
   EY[1] <- X[1];
   Y[1] ~ dnorm(EY[1], inv.r);
   for(i in 2:N) {
      X[i] ~ dnorm(X[i-1] + u, inv.q);
      EY[i] <- X[i];
      Y[i] ~ dnorm(EY[i], inv.r); 
   }
}  
"
jags.paramsss <- c("q", "r", "EY", "u")

jags.datass.psl <- list(Y = as.vector(br_polls2$PSL), N = length(as.vector(br_polls2$PSL)), Y1 = as.vector(br_polls2$PSL)[1])
jags.datass.pt <- list(Y = as.vector(br_polls2$PT), N = length(as.vector(br_polls2$PT)), Y1 = as.vector(br_polls2$PT)[1])
jags.datass.pdt <- list(Y = as.vector(br_polls2$PDT), N = length(as.vector(br_polls2$PDT)), Y1 = as.vector(br_polls2$PDT)[1])
jags.datass.psdb <- list(Y = as.vector(br_polls2$PSDB), N = length(as.vector(br_polls2$PSDB)), Y1 = as.vector(br_polls2$PSDB)[1])
jags.datass.rede <- list(Y = as.vector(br_polls2$REDE), N = length(as.vector(br_polls2$REDE)), Y1 = as.vector(br_polls2$REDE)[1])
jags.datass.others <- list(Y = as.vector(br_polls2$Others), N = length(as.vector(br_polls2$Others)), Y1 = as.vector(br_polls2$Others)[1])

modss.psl = jags.model(textConnection(jagsscriptss), data=jags.datass.psl, n.chains=3)
update(modss.psl, 1e3)
mod_simss.psl = coda.samples(model=modss.psl, variable.names=jags.paramsss, n.iter=5e3)

modss.pt = jags.model(textConnection(jagsscriptss), data=jags.datass.pt, n.chains=3)
update(modss.pt, 1e3)
mod_simss.pt = coda.samples(model=modss.pt, variable.names=jags.paramsss, n.iter=5e3)

modss.pdt = jags.model(textConnection(jagsscriptss), data=jags.datass.pdt, n.chains=3)
update(modss.pdt, 1e3)
mod_simss.pdt = coda.samples(model=modss.pdt, variable.names=jags.paramsss, n.iter=5e3)

modss.psdb = jags.model(textConnection(jagsscriptss), data=jags.datass.psdb, n.chains=3)
update(modss.psdb, 1e3)
mod_simss.psdb = coda.samples(model=modss.psdb, variable.names=jags.paramsss, n.iter=5e3)

modss.rede = jags.model(textConnection(jagsscriptss), data=jags.datass.rede, n.chains=3)
update(modss.rede, 1e3)
mod_simss.rede = coda.samples(model=modss.rede, variable.names=jags.paramsss, n.iter=5e3)

modss.other = jags.model(textConnection(jagsscriptss), data=jags.datass.others, n.chains=3)
update(modss.other, 1e3)
mod_simss.other = coda.samples(model=modss.other, variable.names=jags.paramsss, n.iter=5e3)


mcmcresultspsl = as.mcmc(do.call(rbind, mod_simss.psl))
mcmcresultspt = as.mcmc(do.call(rbind, mod_simss.pt))
mcmcresultspdt = as.mcmc(do.call(rbind, mod_simss.pdt))
mcmcresultspsdb = as.mcmc(do.call(rbind, mod_simss.psdb))
mcmcresultsrede = as.mcmc(do.call(rbind, mod_simss.rede))
mcmcresultsother = as.mcmc(do.call(rbind, mod_simss.other))

(colMeans(round(mcmcresultspsl,2)))
(colMeans(round(mcmcresultspt,2)))
(colMeans(round(mcmcresultspdt,2)))
(colMeans(round(mcmcresultspsdb,2)))
(colMeans(round(mcmcresultsrede,2)))
(colMeans(round(mcmcresultsother,2)))

```